---
title: "Nonstationary"
output: html_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
install.packages("ismev")
install.packages("purrr")
install.packages("extRemes")
library(extRemes)
library(ismev)
library(dplyr)
library(purrr)
```

## Define Functions

We’ll create three different GEV models:

Model 1: Varying loc with scale stationary.
Model 2: Varying scale with loc stationary.
Model 3: Varying both loc and scale.
The function will calculate the BIC for each model and select the model with the lowest BIC.

```{r models}
fit_gev_nonstationary <- function(data, county, variable, debug = FALSE) {  
  subset_data <- data %>% filter(county == !!county)
  t <- as.numeric(subset_data$year)
  ydat <- data.frame(t = t)
  
  # Print the main data being fitted if debug is TRUE
  if (debug) {
    cat("Fitting for county:", county, "variable:", variable, "\n")
    print(subset_data[[variable]])
    print(t)
  }
  
  tryCatch({
    # Model 1: Nonstationary loc, stationary scale
    model_1 <- if (debug) {
      fevd(subset_data[[variable]], data = ydat, location.fun = ~ t, scale.fun = ~ 1, type = "GEV")
    } else {
      suppressMessages(suppressWarnings(
        fevd(subset_data[[variable]], data = ydat, location.fun = ~ t, scale.fun = ~ 1, type = "GEV")
      ))
    }
    if (debug) {
      sum_1 <- summary(model_1)
    } else {
      capture.output(sum_1 <- summary(model_1))
    }
    bic_1 <- sum_1$BIC
    if (debug) cat("Model 1 BIC:", bic_1, "Params:", model_1$results$par, "\n")
    
    # Model 2: Stationary loc, nonstationary scale
    model_2 <- if (debug) {
      fevd(subset_data[[variable]], data = ydat, location.fun = ~ 1, scale.fun = ~ t, type = "GEV")
    } else {
      suppressMessages(suppressWarnings(
        fevd(subset_data[[variable]], data = ydat, location.fun = ~ 1, scale.fun = ~ t, type = "GEV")
      ))
    }
    if (debug) {
      sum_2 <- summary(model_2)
    } else {
      capture.output(sum_2 <- summary(model_2))
    }
    bic_2 <- sum_2$BIC
    if (debug) cat("Model 2 BIC:", bic_2, "Params:", model_2$results$par, "\n")
    
    # Model 3: Nonstationary loc and scale
    model_3 <- if (debug) {
      fevd(subset_data[[variable]], data = ydat, location.fun = ~ t, scale.fun = ~ t, type = "GEV")
    } else {
      suppressMessages(suppressWarnings(
        fevd(subset_data[[variable]], data = ydat, location.fun = ~ t, scale.fun = ~ t, type = "GEV")
      ))
    }
    if (debug) {
      sum_3 <- summary(model_3)
    } else {
      capture.output(sum_3 <- summary(model_3))
    }
    bic_3 <- sum_3$BIC
    if (debug) cat("Model 3 BIC:", bic_3, "Params:", model_3$results$par, "\n")
    
    # Determine the best model based on BIC
    best_model <- which.min(c(bic_1, bic_2, bic_3))
    
    # Function to safely extract parameter if it exists
    get_param <- function(model, param) {
      if (param %in% names(model$results$par)) {
        return(model$results$par[[param]])
      } else {
        return(NA)
      }
    }
    
    # Extract parameters based on the best model
    if (best_model == 1) {
      return(c("type" = 1, "shape" = get_param(model_1, "shape"),
               "a" = get_param(model_1, "scale"), "b" = 0,
               "c" = get_param(model_1, "mu0"), "d" = get_param(model_1, "mu1")))
    } else if (best_model == 2) {
      return(c("type" = 2, "shape" = get_param(model_2, "shape"),
               "a" = get_param(model_2, "sigma0"), "b" = get_param(model_2, "sigma1"),
               "c" = get_param(model_2, "location"), "d" = 0))
    } else {
      return(c("type" = 3, "shape" = get_param(model_3, "shape"),
               "a" = get_param(model_3, "sigma0"),
               "b" = get_param(model_3, "sigma1"),
               "c" = get_param(model_3, "mu0"),
               "d" = get_param(model_3, "mu1")))
    }
  }, error = function(e) {
    # Print error message if debug is TRUE
    if (debug) {
      cat("Error in model fitting for county:", county, "variable:", variable, "\n")
      print(e)
    }
    return(c("type" = NA, "shape" = NA, "a" = NA, "b" = NA, "c" = NA, "d" = NA))
  })
}
```

This function will apply the fit_gev_nonstationary function across each county and each variable, storing the results in a new dataframe.
```{r wrapper}
fit_all_counties <- function(data) {
  counties <- unique(data$county)
  variables <- colnames(data)[!colnames(data) %in% c("year", "county")]
  
  # Initialize an empty list to store all county results
  all_results <- list()
  
  for (county in counties) {
    county_params <- c("county" = county)
    
    for (variable in variables) {
      params <- fit_gev_nonstationary(data, county, variable)
      
      # Concatenate non-NA values for the variable-specific parameters
      for (param_name in names(params)) {
        county_params[[paste0(param_name, "_", variable)]] <- params[[param_name]]
      }
    }
    
    # Append each county's consolidated parameters to the results list
    all_results[[county]] <- as.data.frame(t(county_params), stringsAsFactors = FALSE)
    
    # Print progress message
    cat("Completed county:", county, "\n")
  }
  
  # Bind all county results into a single dataframe
  results <- do.call(rbind, all_results)
  
  # Ensure all columns are consistent and add any missing columns with NA values
  all_columns <- c("county")
  for (variable in variables) {
    all_columns <- c(all_columns, paste0(c("type", "shape", "a", "b", "c", "d"), "_", variable))
  }
  results[setdiff(all_columns, colnames(results))] <- NA
  results <- results[, all_columns]
  
  return(results)
}
```

## Define Test

Test the function

```{r main}
data <- read.csv("MSWEP_Ann_Max/Ann_max_precip_county_MASTER.csv",
  colClasses = c(county = "character", year = "numeric"))

# Choose a single county and variable for testing
test_county <- data$county[1]  # Select the first county in the dataset
test_variable <- colnames(data)[3]  # Select the first variable column (after 'county' and 'year')

# Print to verify
cat("Testing with county:", test_county, "and variable:", test_variable, "\n")

# Run the function for one county and one variable
test_result <- fit_gev_nonstationary(data, test_county, test_variable, debug=TRUE)

# Display the test result
print(test_result)
```

## Define Main

Load the data, call the function, and save the results. 

For MSWEP:

```{r main_mswep}
# Load your dataset
data <- read.csv("MSWEP_Ann_Max/Ann_max_precip_county_MASTER.csv",
  colClasses = c(county = "character", year = "numeric"))

# Fit the models and get parameters
param_results <- fit_all_counties(data)

# Save the results to CSV
write.csv(param_results, "MSWEP_Params.csv", row.names = FALSE)
```

For ERA5:

```{r main_era5}
# Load your dataset
data <- read.csv("ERA5_Ann_Max/ERA5_Ann_max_precip_county_MASTER.csv",
  colClasses = c(county = "character", year = "numeric"))

# Fit the models and get parameters
param_results <- fit_all_counties(data)

# Save the results to CSV
write.csv(param_results, "ERA5_Params.csv", row.names = FALSE)
```
